### 2017-01-06 experiments.tex
version submitted to sir

\begin{table}
\centering
 \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-4thf}
\scalebox{0.8}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
&&\\
\textbf{2SPSA-3} & $0.080 \pm 0.0053$ & $0.043 \pm 0.0015$\\
&&\\
\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
&&\\
\textbf{2SPSA-3} & $0.0338 \pm 0.0001$ & $0.0315 \pm 0.0007$\\
&&\\
\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
 \bottomrule
\end{tabular}
}
\end{table}

@@

 Figure \ref{fig:norlossvssims} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$. Figure \ref{fig:QuadnorlossvssimsNoiseless} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$. See Figures \ref{fig:norlossvssimsNoiseless}--\ref{fig:norlossvssimsNoiseless}  for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss. 
 From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Fig \ref{fig:norlossvssims}, we make the following observations:

 \textit{\textbf{Observation 1:} Among 2RDSA schemes, 2RDSA-IH performs better than regular 2RDSA, for both perturbation choices.}

\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH, with 2RDSA-IH-AsymBer performing the best overall.}

>>>>

\begin{table}
\centering
 \caption{Normalized loss values for fourth-order  objective \eqref{eq:4thorder} with and without noise: standard error from $500$ replications shown after $\pm$}
\label{tab:norloss-4thf}
\scalebox{0.8}{
\begin{tabular}{|c|c|c|}
\toprule
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0.1$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.132 \pm 0.0267$ & $0.104 \pm 0.0355$\\
&&\\
\textbf{2SPSA-3} & $0.0893 \pm 0.001$ & $0.0457 \pm 0.0004$\\
&&\\
\textbf{2RDSA-Unif} &$0.115 \pm 0.0214$ & $0.0271 \pm 0.0538$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0471 \pm 0.021$& $\bm{0.0099 \pm 0.0014}$\\
 \bottomrule
 %%%%%%%%%%%%2nd ordr
\rowcolor{gray!20}
\multicolumn{3}{||c|}{\multirow{2}{*}{\textbf{Noise parameter $\sigma=0$}}}\\[1em]
%&&&\\
\midrule
  & \multirow{2}{*}{\textbf{Regular}} & \textbf{Improved Hessian}  \\
  & & \textbf{ estimation} \\
 \midrule
\textbf{2SPSA} & $0.0795 \pm 0.0234$ & $0.0628 \pm 0.0234$\\
&&\\
\textbf{2SPSA-3} & $0.0338 \pm 0.0001$ & $0.0315 \pm 0.0007$\\
&&\\
\textbf{2RDSA-Unif} &$0.0813 \pm 0.0275$ & $0.0214 \pm 0.00376$\\ 
&&\\
\textbf{2RDSA-AsymBer}& $0.0199 \pm 0.0114$& $\bm{0.0098 \pm 0.00147}$\\
 \bottomrule
\end{tabular}
}
\end{table}

@@

 Figures \ref{fig:norlossvssims}, \ref{fig:norlossvssimsspsa-3} plots the normalized loss as a function of the simulation budget with the fourth-order loss objective with $\sigma=0.1$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA. Figures \ref{fig:QuadnorlossvssimsNoiseless}, \ref{fig:QuadnorlossvssimsNoiselessspsa3} plots the normalized loss as a function of the simulation budget with the quadratic loss objective with $\sigma=0$ for algorithms 2RDSA and 2SPSA-3 respectively when compared with 2SPSA. See Figures \ref{fig:norlossvssimsNoiseless}-- \ref{fig:QuadnorlossvssimsNoisy} and \ref{fig:norlossvssimsNoiselessspsa3}-- \ref{fig:QuadnorlossvssimsNoisyspsa3} for similar results with $\sigma = 0$ for fourth-order loss and $\sigma=0.1$ for quadratic loss. 
 From the results in Tables \ref{tab:norloss-4thf}--\ref{tab:nmse-quadratic} and Figures, we make the following observations:

 \textit{\textbf{Observation 1:} Schemes with improved Hessian estimation performs better than their respective regular schemes.}

\textit{\textbf{Observation 2:} 2RDSA-IH variants outperform both 2SPSA and 2SPSA-IH.}

\textit{\textbf{Observation 3:} 2SPSA-3-IH variants outperform both 2SPSA and 2SPSA-IH.}

\textit{\textbf{Observation 4:} 2RDSA-IH-AsymBer performing the best overall.}

\begin{figure}
    \centering
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{5}{6}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};

      % 2SPSA-3
      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};
			%%% With IH
      % 2SPSA-3
      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations_noisycase.csv}{0}{7}{8}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {norloss_vs_simulations_noisycase_spsa3.csv};

            \end{axis}
      \end{tikzpicture}}\\}
			\caption{Normalized loss vs. number of simulations for fourth-order loss \eqref{eq:4thorder} with $\sigma=0.1$ for 2SPSA, 2SPSA-3 algorithms with/without improved Hessian estimation: bands around the curves represent standard error from $500$ replications. }
      \label{fig:norlossvssimsspsa-3} 
\end{figure}

%%%%%% Norloss-Quadratic-noisefree %%%%%%%%%%%%%%
\begin{figure}
    \centering
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH  },
        %%legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_spsa3.csv};
			%%%% With IH
      %2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{5}{6}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma] {quad_norloss_vs_simulations_spsa3.csv};
			%
      % 2SPSA-3
      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_spsa3.csv};
			%%%% With IH
      % 2SPSA-3
      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations.csv}{0}{7}{8}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma]  {quad_norloss_vs_simulations_spsa3.csv};

           \end{axis}
      \end{tikzpicture}}\\}
			\caption{Normalized loss vs. number of simulations for quadratic loss \eqref{eq:quadratic} with $\sigma=0$ for 2SPSA, 2SPSA-3 algorithms with/without improved Hessian estimation. }
      \label{fig:QuadnorlossvssimsNoiselessspsa3} 
\end{figure}

\begin{figure}[h]
    \centering
		\begin{tabular}{c}
		\begin{subfigure}{0.5\textwidth}
\tabl{c}{\scalebox{0.8}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {norloss_vs_simulations_spsa3.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{5}{6}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma]  {norloss_vs_simulations_spsa3.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {norloss_vs_simulations_spsa3.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{norloss_vs_simulations.csv}{0}{7}{8}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma]  {norloss_vs_simulations_spsa3.csv};

            \end{axis}
      \end{tikzpicture}}\\}
			\caption{Fourth-order loss \eqref{eq:4thorder} with $\sigma=0$. }
      \label{fig:norlossvssimsNoiselessspsa3} 
\end{subfigure} 
\\
\\
%%%%%%% Norloss-Quadratic-noisy %%%%%%%%%%%%%%
\begin{subfigure}{0.5\textwidth}
\tabl{c}{\scalebox{0.9}{\begin{tikzpicture}
      \begin{axis}[
	xlabel={number of simulations},
	ylabel={Normalized loss},
       legend entries={
	 ,2SPSA,,2SPSA-IH,,2SPSA-3,,2SPSA-3-IH },
        %legend pos=outer north east,
			legend style={
at={(0,0)},
anchor=north,at={(axis description cs:0.5,-0.2)},legend columns=2}	]
      % 2SPSA
      \errorband[red!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{1}{2}
      \addplot [thick, red,mark=square*,mark options={scale=0.7}] table [x index=0, y index=1,col sep=comma] {quad_norloss_vs_simulations_noisycase_spsa3.csv};
			%%% With IH
      % 2SPSA
      \errorband[darkgray!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{5}{6}
      \addplot [thick, darkgray,mark=*,mark options={scale=0.75}] table [x index=0, y index=5,col sep=comma]  {quad_norloss_vs_simulations_noisycase_spsa3.csv};

      % 2RDSA-Unif
      \errorband[blue!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{3}{4}
      \addplot [thick, blue,mark=square*,mark options={scale=0.7}] table [x index=0, y index=3,col sep=comma] {quad_norloss_vs_simulations_noisycase_spsa3.csv};
			%%% With IH
      % 2RDSA-Unif
      \errorband[magenta!50!white, opacity=0.3]{quad_norloss_vs_simulations_noisycase.csv}{0}{7}{8}
      \addplot [thick, magenta,mark=*,mark options={scale=0.75}] table [x index=0, y index=7,col sep=comma] {quad_norloss_vs_simulations_noisycase_spsa3.csv};

            \end{axis}
      \end{tikzpicture}}\\}
			\caption{Quadratic loss \eqref{eq:quadratic} with $\sigma=0.1$. }
      \label{fig:QuadnorlossvssimsNoisyspsa3} 
\end{subfigure} 
		\end{tabular}
			\caption{Normalized loss vs. number of simulations in two different loss settings for  2SPSA, 2SPSA-3 algorithms. }
      \label{fig:extras} 
		\end{figure}

### 2017-01-06 intro.tex
version submitted to sir

We establish that the proposed improvements to Hessian estimation in 2RDSA are such  that the resulting 2RDSA-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2RDSA-IH converges almost surely to the true Hessian. 
%Moreover, we show, for the special case of a quadratic objective, that	2RDSA-IH results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. 
Further, we show empirically that 2RDSA-IH outperforms both 2SPSA-IH of \cite{spall-jacobian} and regular 2RDSA of \cite{prashanth2015rdsa}. Our contribution is important because 2RDSA-IH, like 2RDSA, has lower simulation cost per iteration than 2SPSA and unlike 2RDSA, has an improved Hessian estimation scheme.
\item Our work in chapter  \ref{sec:2rdsa-gen} is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}. In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions. In chapter \ref{sec:2rdsa-gen} , we proposed gradient and Hessian estimation schemes which are independent of the perturbation distributions. However, perturbations have to satisfy the i.i.d, mean-zero assumptions. Apart from these common assumptions  one additional assumption like difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required. This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions. Some well known distribution that can used for these schemes as a result of generalisation are Normal (0,1) and Cauchy distributions.
\item Our work in \ref{sec:2spsa-3-ih} is centred on improving the second order SPSA scheme with three simulations ( referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}. This improvements are as a result of incorporating the zero mean feedback term in Hessian estimate and optimizing the step-sizes used in the Hessian estimate recursion. Though these ideas are same as the ideas presented in this section for 2RDSA scheme, the contribution is non-trivial because the derivation of feedback term and Hessian estimate is entirely different from 2SPSA, 2RDSA improvements. Moreover, we show, for the special case of a quadratic objective and when there is no noise, that 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.
\end{enumerate}

>>>>

We establish that the proposed improvements to Hessian estimation in 2RDSA are such  that the resulting 2RDSA-IH algorithm is provably convergent, in particular, the Hessian estimate $\overline H_n$ of 2RDSA-IH converges almost surely to the true Hessian. 
%Moreover, we show, for the special case of a quadratic objective, that	2RDSA-IH results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. 
Further, we show empirically that 2RDSA-IH outperforms both 2SPSA-IH of \cite{spall-jacobian} and regular 2RDSA of \cite{prashanth2015rdsa}. Our contribution is important because 2RDSA-IH, like 2RDSA, has lower simulation cost per iteration than 2SPSA and unlike 2RDSA, has an improved Hessian estimation scheme.
\item Our work in chapter  \ref{sec:2rdsa-gen} is centred on generalising the 2RDSA scheme of \cite{prashanth2015rdsa}. In \cite{prashanth2015rdsa}, 2RDSA scheme involving only two perturbation distributions was proposed, those are uniform and asymmetric Bernoulli distributions. In chapter \ref{sec:2rdsa-gen} , we proposed gradient and Hessian estimation schemes which are independent of the perturbation distributions. However, perturbations have to satisfy the i.i.d, mean-zero assumptions. Apart from these common assumptions  one additional assumption, which is the difference between the second moment square  and fourth moment should be non zero for Hessian estimation is also required. This generalisation of distributions of random variables for perturbations is important because it enhances the scope of improving the performance of the 1RDSA and 2RDSA schemes by incorporating several other distributions. Some well known distribution that can used for these schemes as a result of generalisation are Normal (0,1) and Cauchy distributions.
\item Our work in \ref{sec:2spsa-3-ih} is centred on improving the second order SPSA scheme with three simulations ( referred to as 2SPSA-3 hereafter), which was proposed in  \cite{bhatnagar2015simultaneous}. This improvements are as a result of incorporating the zero mean feedback term in Hessian estimate and optimizing the step-sizes used in the Hessian estimate recursion. Though these ideas are same as the ideas presented in this section for 2RDSA scheme, the contribution is non-trivial because the derivation of feedback term and Hessian estimate is entirely different from 2SPSA, 2RDSA improvements. Moreover, we show, for the special case of a quadratic objective and when there is no noise, that 2SPSA-3 scheme along with proposed improvement to Hessian estimation (henceforth referred to as 2SPSA-3-IH) results in a convergence rate that is on par with the corresponding rate for 2SPSA with Hessian estimation improvements (2SPSA-IH) \cite{spall-jacobian}. The advantage with 2SPSA-3-IH is that it requires only 75\% of the simulation cost per-iteration for 2SPSA-IH.
\end{enumerate}
